Measuring individual differences in the understanding of gaze cues across the lifespan

### 2022-08-09 paper/manuscript.Rmd
save last state before dropping vali section

# Validity
Our third aim was to assess whether the captured individual variation in gaze understanding relates to factors in children's real live social environment. Previous studies found associations between social cognition measures and various environmental factors [@devine2018family; @hughes2004links], including family background and education [@cutting1999theory; bulgarelli2016social], number and age of siblings and family constellation [@zhang2021worse, @cassidy2005theory; @perner1994theory; @peterson2000kindred], interaction with siblings [@dunn1991young], and centre-based childcare [@bulgarelli2016social]. It is assumed that opportunities to play, communicate and argue with siblings and similarly-aged peers help children to understand the human mind. Therefore, if we find a link between gaze understanding and family factors, we regard this as an indicator for the predictive validity of our measure.

@@

```{r fig4, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab4)"}
knitr::include_graphics("../figures/extvali_results.png")
```

@@

# Discussion
We presented a new experimental paradigm to study individual differences in a fundamental social-cognitive ability -- gaze understanding. 
Participants are asked to locate a balloon by using the gaze cues of another agent. 
The outcome measure depends on the presented study version: 
in the hedge version, we implemented a continuous measure, while the box version employs a categorical measure. 
The cartoon-like presentation makes the task engaging for children and adults alike.

- summary (so als ob die leute nix gelesen hätten)
  - continuous
  - across age, adults nochmal aufgreifen
  - good reliability

## Comparing our study versions
Our continuous hedge version yields higher internal consistency estimates compared to the categorical box version. 
Both study versions exhibit high retest reliability. 
Therefore, when a sufficient amount of trials is presented, the box version of the task can yield reliable individual estimates [cf. @hughes2000good improved reliability through aggregation]. 
With respect to psychometric properties, we would recommend the use of the continuous version. 
However, the categorical box version demonstrates design features that might be preferable in some research contexts: the discrete outcome might facilitate clicking for children. In addition, researchers could induce different levels of salience for each box. Our task could consequently be used to study bias, preferences and diverse desires (e.g. matching the box appearance to some feature of the agent).

## Online implementation
For our task, we designed a new testing infrastructure. 
The gaze understanding task is presented as an interactive web app. 
This enables presentation across devices without any prior installation. 
Short trials facilitate a substantial number of trials -- here, we employed 15 test trials within 5-10 minutes testing time.
Consequently, individual-level estimates are more precise. 
Stimuli presentation is achieved through the use of SVGs. This has several advantages: 
the aspect ratio and stimulus quality are kept constant no matter which size the web browser displays. 
Most importantly, we can dynamically modify the stimulus details (e.g., positions) on a trial-by-trial basis.
Presented agents and objects can be easily adapted for future task modifications or specific cultural settings.

## Comparing data collection modes
Participants can be tested in-person with supervision or remotely at home. 
Test instructions are standardized and, with prior informed consent, the webcam records study participation.
This allows us to scale up data collection: testing is flexible, fast and requires no further experimenter training. 
We compared children participating in-person, supervised in kindergartens with children participated remotely at home. 
Our results suggest a comparable developmental trajectory of gaze understanding in both samples. 
Children in the remote sample are slightly more precise. 
This effect is most pronounced in the three-year-olds in the box version (for an analysis of the webcam recordings, see Appendix XXX). 
Therefore, we recommend using a tablet for remote data collection. 
Children can then click for themselves and caregivers have less chance to interfere.

## Validity
Validity is often assessed by looking at concurrent relations between measures. Here we looked at a relation that is often postulated for social cognition tasks, namely that social-cognitive ability is predicted by family level variables.
We found that children's gaze understanding relates to factors of their social environment: 
children were more accurate, the more social exposure and the more diverse their surrounding was.
However, compared to our fine-grained, parametric measure of gaze understanding, family variables were only captured on the surface level by employing single questionnaire items. 
We hope that future technological advances will enable continuous measures of children's real life social surroundings [@barr2020screen; @long2020detecting, @rogoff2018importance].

@@

- conclusion:
  - big picture
  - ind diff important for ..., good measures needed
  - dedicated measures of ind diff instead of simply using well-established. reliability    paradox paper zitieren
  - task & infrastructure
  - foundations for more precise measures for cognitive development

>>>>

# Validity
Our third aim was to assess whether the captured individual variation in gaze understanding relates to factors in children's real live social environment. Previous studies found associations between social cognition measures and various environmental factors [@devine2018family; @hughes2004links], including family background and education [@cutting1999theory; bulgarelli2016social], number and age of siblings and family constellation [@zhang2021worse, @cassidy2005theory; @perner1994theory; @peterson2000kindred], interaction with siblings [@dunn1991young], and centre-based childcare [@bulgarelli2016social]. It is assumed that opportunities to play, communicate and argue with siblings and similarly-aged peers help children to understand the human mind. Therefore, if we find a link between gaze understanding and family factors, we regard this as an indicator for the validity of our measure.

@@

```{r fig4, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab4)", fig.show='asis'}
knitr::include_graphics("../figures/extvali_results.png")
```

@@

# General Discussion
In this paper, we presented a new experimental paradigm to study individual differences in gaze understanding across the lifespan. 
This paper contributes to methodological advances in developmental psychology in the following ways. 
First, our new web-based testing infrastructure proved useful for standardized, portable data collection at scale, both remotely as well as in-person. 
Second, we could captured fine-grained individual differences in gaze understanding at different ages -- from early childhood until adulthood.
Third, our task showed satisfactory psychometric properties with respect to internal consistency and retest reliability estimates. 
Finally, we found that gaze understanding relates to aspects of children's everyday experience.
The web app (https://ccp-odc.eva.mpg.de/gafo-demo/) and its source code (https://github.com/ccp-eva/gafo-demo) are openly accessible.
We want to highlight that researchers are welcome to use and modify our task according to their needs.

One of the main contributions of this paper is that our newly presented gaze understanding task induces variation between participants. 
This is primarily achieved through the implementation of short trials, resulting in over a dozen replicates per subject (15 test trials within 5-10 minutes testing time).
Our experimental design enables a continuous outcome measure, which again quantifies individual differences.

In our task, participants are asked to locate a balloon by using the gaze cues of another agent. 
The outcome measure depends on the presented study version: 
in the hedge version, we implemented a continuous measure, while the box version employs a categorical measure. 
Our continuous hedge version yields higher internal consistency estimates compared to the categorical box version. 
Both study versions exhibit high retest reliability. 
Therefore, when a sufficient amount of trials is presented, the box version of the task can yield reliable individual estimates [cf. @hughes2000good; improved reliability through aggregation]. 
When testing time is limited, we would recommend the use of the continuous study version for higher internal consistency. 
However, the categorical box version demonstrates design features that might be preferable in some research contexts: the discrete outcome might facilitate clicking for children. In addition, researchers could induce different levels of salience for each box. Our task could consequently be used to study bias, preferences and diverse desires (e.g. matching the box appearance to some feature of the agent). 
In sum, our task provides a step towards more robust and reliable research methods, especially with regards to individual differences research.

In addition to the new task design itself, we designed a new testing infrastructure. 
Our gaze understanding task is presented as an interactive web app. 
This enables presentation across devices without any prior installation. 
Stimuli presentation is achieved through the use of SVGs. This has several advantages: 
the aspect ratio and stimulus quality are kept constant no matter which size the web browser displays. 
Most importantly, we can dynamically modify the stimulus details (e.g., positions) on a trial-by-trial basis.
Presented agents and objects can be easily adapted for future task modifications or specific cultural settings. 
The cartoon-like presentation makes the task engaging for children and adults alike.

Participants can be tested in-person with supervision or remotely at home. 
Test instructions are standardized and, with prior informed consent, the webcam records study participation.
This allows us to scale up data collection: testing is flexible, fast and requires no further experimenter training. 
We compared children participating in-person, supervised in kindergartens with children participated remotely at home. 
Our results suggest a comparable developmental trajectory of gaze understanding in both samples. 
Children in the remote sample are slightly more precise. 
This effect is most pronounced in the three-year-olds in the box version (for an analysis of the webcam recordings, see Appendix XXX). 
Therefore, we recommend using a tablet for remote data collection. 
Children can then click for themselves and caregivers have less chance to interfere.

After finding consistent and reliable individual estimates, we aimed at verifying the appropriateness of our new task.
Validity is often assessed by looking at concurrent relations between measures. Here we looked at a relation that is often postulated for social cognition tasks, namely that social-cognitive ability is predicted by family level variables.
We found that children's gaze understanding relates to factors of their social environment: 
children were more accurate, the more social exposure and the more diverse their surrounding was.
However, compared to our fine-grained, parametric measure of gaze understanding, family variables were only captured on the surface level by employing single questionnaire items. 
We hope that future technological advances will enable continuous measures of children's real life social surroundings [@barr2020screen; @long2020detecting; @rogoff2018importance].

@@

# Limitations

# Conclusion
  - big picture
  - ind diff important for ..., good measures needed
  - dedicated measures of ind diff instead of simply using well-established. reliability    paradox paper zitieren
  - task & infrastructure
  - foundations for more precise measures for cognitive development

### 2022-08-09 paper/manuscript.tex
save last state before dropping vali section

Our third aim was to assess whether the captured individual variation in gaze understanding relates to factors in children's real live social environment. Previous studies found associations between social cognition measures and various environmental factors (Devine \& Hughes, 2018; Hughes \& Leekam, 2004), including family background and education {[}Cutting and Dunn (1999); bulgarelli2016social{]}, number and age of siblings and family constellation Cassidy, Fineberg, Brown, \& Perkins (2005), interaction with siblings (Dunn et al., 1991), and centre-based childcare (Bulgarelli \& Molina, 2016). It is assumed that opportunities to play, communicate and argue with siblings and similarly-aged peers help children to understand the human mind. Therefore, if we find a link between gaze understanding and family factors, we regard this as an indicator for the predictive validity of our measure.

@@

\hypertarget{discussion-3}{%
\section{Discussion}\label{discussion-3}}

We presented a new experimental paradigm to study individual differences in a fundamental social-cognitive ability -- gaze understanding.
Participants are asked to locate a balloon by using the gaze cues of another agent.
The outcome measure depends on the presented study version:
in the hedge version, we implemented a continuous measure, while the box version employs a categorical measure.
The cartoon-like presentation makes the task engaging for children and adults alike.

\begin{itemize}
\tightlist
\item
  summary (so als ob die leute nix gelesen hätten)

  \begin{itemize}
  \tightlist
  \item
    continuous
  \item
    across age, adults nochmal aufgreifen
  \item
    good reliability
  \end{itemize}
\end{itemize}

\hypertarget{comparing-our-study-versions}{%
\subsection{Comparing our study versions}\label{comparing-our-study-versions}}

Our continuous hedge version yields higher internal consistency estimates compared to the categorical box version.
Both study versions exhibit high retest reliability.
Therefore, when a sufficient amount of trials is presented, the box version of the task can yield reliable individual estimates (cf. Hughes et al., 2000 improved reliability through aggregation).
With respect to psychometric properties, we would recommend the use of the continuous version.
However, the categorical box version demonstrates design features that might be preferable in some research contexts: the discrete outcome might facilitate clicking for children. In addition, researchers could induce different levels of salience for each box. Our task could consequently be used to study bias, preferences and diverse desires (e.g.~matching the box appearance to some feature of the agent).

\hypertarget{online-implementation}{%
\subsection{Online implementation}\label{online-implementation}}

For our task, we designed a new testing infrastructure.
The gaze understanding task is presented as an interactive web app.
This enables presentation across devices without any prior installation.
Short trials facilitate a substantial number of trials -- here, we employed 15 test trials within 5-10 minutes testing time.
Consequently, individual-level estimates are more precise.
Stimuli presentation is achieved through the use of SVGs. This has several advantages:
the aspect ratio and stimulus quality are kept constant no matter which size the web browser displays.
Most importantly, we can dynamically modify the stimulus details (e.g., positions) on a trial-by-trial basis.
Presented agents and objects can be easily adapted for future task modifications or specific cultural settings.

\hypertarget{comparing-data-collection-modes}{%
\subsection{Comparing data collection modes}\label{comparing-data-collection-modes}}

@@

\hypertarget{validity-1}{%
\subsection{Validity}\label{validity-1}}

Validity is often assessed by looking at concurrent relations between measures. Here we looked at a relation that is often postulated for social cognition tasks, namely that social-cognitive ability is predicted by family level variables.
We found that children's gaze understanding relates to factors of their social environment:
children were more accurate, the more social exposure and the more diverse their surrounding was.
However, compared to our fine-grained, parametric measure of gaze understanding, family variables were only captured on the surface level by employing single questionnaire items.
We hope that future technological advances will enable continuous measures of children's real life social surroundings Rogoff, Dahl, \& Callanan (2018).

@@

\begin{itemize}
\tightlist
\item
  conclusion:

  \begin{itemize}
  \tightlist
  \item
    big picture
  \item
    ind diff important for \ldots, good measures needed
  \item
    dedicated measures of ind diff instead of simply using well-established. reliability paradox paper zitieren
  \item
    task \& infrastructure
  \item
    foundations for more precise measures for cognitive development
  \end{itemize}
\end{itemize}

>>>>

Our third aim was to assess whether the captured individual variation in gaze understanding relates to factors in children's real live social environment. Previous studies found associations between social cognition measures and various environmental factors (Devine \& Hughes, 2018; Hughes \& Leekam, 2004), including family background and education {[}Cutting and Dunn (1999); bulgarelli2016social{]}, number and age of siblings and family constellation Cassidy, Fineberg, Brown, \& Perkins (2005), interaction with siblings (Dunn et al., 1991), and centre-based childcare (Bulgarelli \& Molina, 2016). It is assumed that opportunities to play, communicate and argue with siblings and similarly-aged peers help children to understand the human mind. Therefore, if we find a link between gaze understanding and family factors, we regard this as an indicator for the validity of our measure.

@@

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

In this paper, we presented a new experimental paradigm to study individual differences in gaze understanding across the lifespan.
This paper contributes to methodological advances in developmental psychology in the following ways.
First, our new web-based testing infrastructure proved useful for standardized, portable data collection at scale, both remotely as well as in-person.
Second, we could captured fine-grained individual differences in gaze understanding at different ages -- from early childhood until adulthood.
Third, our task showed satisfactory psychometric properties with respect to internal consistency and retest reliability estimates.
Finally, we found that gaze understanding relates to aspects of children's everyday experience.
The web app (\url{https://ccp-odc.eva.mpg.de/gafo-demo/}) and its source code (\url{https://github.com/ccp-eva/gafo-demo}) are openly accessible.
We want to highlight that researchers are welcome to use and modify our task according to their needs.

One of the main contributions of this paper is that our newly presented gaze understanding task induces variation between participants.
This is primarily achieved through the implementation of short trials, resulting in over a dozen replicates per subject (15 test trials within 5-10 minutes testing time).
Our experimental design enables a continuous outcome measure, which again quantifies individual differences.

In our task, participants are asked to locate a balloon by using the gaze cues of another agent.
The outcome measure depends on the presented study version:
in the hedge version, we implemented a continuous measure, while the box version employs a categorical measure.
Our continuous hedge version yields higher internal consistency estimates compared to the categorical box version.
Both study versions exhibit high retest reliability.
Therefore, when a sufficient amount of trials is presented, the box version of the task can yield reliable individual estimates {[}cf. Hughes et al. (2000); improved reliability through aggregation{]}.
When testing time is limited, we would recommend the use of the continuous study version for higher internal consistency.
However, the categorical box version demonstrates design features that might be preferable in some research contexts: the discrete outcome might facilitate clicking for children. In addition, researchers could induce different levels of salience for each box. Our task could consequently be used to study bias, preferences and diverse desires (e.g.~matching the box appearance to some feature of the agent).
In sum, our task provides a step towards more robust and reliable research methods, especially with regards to individual differences research.

In addition to the new task design itself, we designed a new testing infrastructure.
Our gaze understanding task is presented as an interactive web app.
This enables presentation across devices without any prior installation.
Stimuli presentation is achieved through the use of SVGs. This has several advantages:
the aspect ratio and stimulus quality are kept constant no matter which size the web browser displays.
Most importantly, we can dynamically modify the stimulus details (e.g., positions) on a trial-by-trial basis.
Presented agents and objects can be easily adapted for future task modifications or specific cultural settings.
The cartoon-like presentation makes the task engaging for children and adults alike.

@@

After finding consistent and reliable individual estimates, we aimed at verifying the appropriateness of our new task.
Validity is often assessed by looking at concurrent relations between measures. Here we looked at a relation that is often postulated for social cognition tasks, namely that social-cognitive ability is predicted by family level variables.
We found that children's gaze understanding relates to factors of their social environment:
children were more accurate, the more social exposure and the more diverse their surrounding was.
However, compared to our fine-grained, parametric measure of gaze understanding, family variables were only captured on the surface level by employing single questionnaire items.
We hope that future technological advances will enable continuous measures of children's real life social surroundings (Barr et al., 2020; Long, Kachergis, Agrawal, \& Frank, 2020; Rogoff, Dahl, \& Callanan, 2018).

@@

\hypertarget{limitations}{%
\section{Limitations}\label{limitations}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\begin{itemize}
\tightlist
\item
  big picture
\item
  ind diff important for \ldots, good measures needed
\item
  dedicated measures of ind diff instead of simply using well-established. reliability paradox paper zitieren
\item
  task \& infrastructure
\item
  foundations for more precise measures for cognitive development
\end{itemize}
